---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# checker

<!-- badges: start -->
<!-- badges: end -->

One of the challenges with teaching R is that some students come to practicals with very old versions of R, or without critical packages installed, despite instructions.

The `checker` package check whether the recommended version of R, RStudio and packages are installed. 
Information on what versions are required is specified in yaml file. 
A yaml file is included in the installation, and can also be supplied with a URL.

## Installation

You can install the development version of checker from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("richardjtelford/checker")
```

## Usage

The function `chk_requirements()` is the only function the user needs to run.

```{r example}
library(checker)
chk_requirements()
```


