---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# checker

<!-- badges: start -->
[![R-CMD-check](https://github.com/richardjtelford/checker/workflows/R-CMD-check/badge.svg)](https://github.com/richardjtelford/checker/actions)
<!-- badges: end -->

One of the challenges with teaching R is that some students come to practicals with very old versions of R, or without critical packages installed.
There are several solutions to this:

- use [rstudio.cloud](rstudio.cloud), where the instructor can control the versions of software and packages used
- have the class install a package that has all necessary packages for the course as dependencies
- use the `renv` package to install packages listed in a lockfile
- use a [rocker container](https://rocker-project.org/)

One problem with these options is that they take control of package installation, rather than giving the student practice installing packages.

The `checker` attempts to be an alternative solution. 
It checks whether the recommended (or more recent) versions of R, RStudio and packages, as specified in a yaml file, are installed and that recommended RStudio options are set.
A sample yaml file is included in the installation. 
One can also be supplied with a URL or path.

## Installation

You can install the development version of `checker` from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("richardjtelford/checker")
```

## Usage

The function `chk_requirements()` is the only function the user needs to run.

```{r example}
library(checker)
chk_requirements()
```
By default, `chk_requirements()` uses a yaml file included in the installation.
To run `chk_requirements()` with your own set of requirements, you can use a URL or file path.

```{r}
#| eval: false
chk_requirements(path = url("https://raw.githubusercontent.com/richardjtelford/checker/main/inst/default.yaml"))
```


## The yaml file

Below is the yaml file included in the installation. It can be edited to meet your requirements by hand or with the function `chk_make()` which takes data.frames of recommended programs, packages, and RStudio options as arguments. Recommended and/or minimum version and any message can be included in these data.frames.



```{r}
#| echo: FALSE
#| comment: ""
readLines("inst/default.yaml") |> writeLines()
```

The accepted keys are

- r_version
- rstudio
- quarto
- git
- packages

All keys are optional.
The first four take "recommended" and "minimum" to specify the recommended and minimum versions (if only one of "recommended" and "minimum" is set, they are treated in the same). If any version is acceptable, use NA.

The rstudio key also accepts an options field, which takes the name with value set to the recommended value and an optional message. A list of Rstudio options can be found with `usethis:::rstudio_prefs_read()`.

The "packages" key has an element for each package installed. These take the same recommended" and "minimum" fields as above, and also a "message" field which is printed in the package is not installed (the ">" lets the message span several lines). The message could be used to point to the location of packages not available on CRAN.

The dashes denote the start and end of the yaml are optional. The formatting with white space must be followed.
